---
title: "R Notebook"
output: 
  github_document: default

---


We load some packages useful for our analysis.


```{r message = FALSE}

library(reshape2)
library(ggplot2)
library(stringr)
library(xtable)
library(plyr)
library(dplyr)
library(readr)
library(gridExtra)

```


## Importation

We import the data. The first column is duplicated. We delete it.


```{r message = FALSE, warning = FALSE}
data <- read_csv("data.csv")
data <- data[,-1]

```


## What are the variables in the database?


```{r}
str(data)

```

| Variable name | Description |
|----------|-------------|
| ID | ID number in the HistoriaGames database |
| name | Name of the game |
| genre | The game «genre» |
| editor | The game editor, the one who published the game |
| dev | The developer, the one who created the game |
| year | The release date |
| platform | The platform on which the game is available |
| period | The historical period |
| link | URL to the game page on the HistoriaGames website |
| addon | Is it a stand-alone game or an add-on? |


**Note:** the database we received had to be pre-processed in many ways. One game would appear on one line, even if it had more than one platform or more than one historical period. We also had to create the addon variable. The idea was to have a data frame as tidy as possible.


## How many entries in this database?


```{r}

length(unique(data$ID))

```


## How many stand-alone games in this database?


```{r}

length(unique(data[data$addon == "games",]$ID))

```


## How many DLCs in this database?


```{r}

length(unique(data[data$addon == "addon",]$ID))

```


## On how many different platforms is a game available on average?


```{r}

nrow(unique(data[,c(1,7)])) / length(unique(data$ID))

```


## On how many different platforms is a stand-alone game available on average?


```{r}

nrow(unique(data[data$addon == "games",c(1,7)])) / length(unique(data[data$addon == "games",]$ID))

```


## On how many different platforms is a DLC available on average?


```{r}

nrow(unique(data[data$addon == "addon",c(1,7)])) / length(unique(data[data$addon == "addon",]$ID))

```


## How many different platforms are there in total?


```{r}

length(unique(data$platform))

```


## Figure 1a : Number of games per year


```{r}

## We copy the data because we will do some processing
data1 <- data

## We turn the "addon" variable which is a string into a factor.
data1$addon <- factor(data1$addon, levels = c("addon", "games"))

## "addon" are now called "Add-ons"
levels(data1$addon)[levels(data1$addon)=="addon"] <- "Add-ons"

## "games" are now called "Games"
levels(data1$addon)[levels(data1$addon)=="games"] <- "Games"

## We keep only one copy of each game
## The informations we need are ID, year and addon
data1 <- data1[!duplicated(data1$ID),]

```


## This the drawing of fig 1a.


```{r}

p1 <- ggplot(data1, aes(x=year, 
                        fill=addon)) + 
  geom_bar() + 
  scale_fill_grey(start = 0.6, 
                  end = 0.2, 
                  name = element_blank(), 
                  guide = guide_legend(reverse = TRUE)) + 
  scale_x_continuous(breaks = 1981:2015, 
                     minor_breaks = NULL) +
  scale_y_continuous(breaks = (0:5) * 20, 
                     minor_breaks = NULL) +
  xlab("") +
  ylab("Number of historical videogames\n ") + 
  theme(axis.text.x = element_text(angle = 90, 
                                   hjust = 1, 
                                   vjust = 0.5)) + 
  theme(legend.position = "bottom") +
  theme(panel.background = element_rect(fill = "grey92")) +
  theme(axis.ticks = element_blank())

p1

ggsave("fig_1a.png", width = 10)

```


## The number of games released every year


```{r}

table(unique(data[c("ID", "year")])$year)

```


## A function giving the list of games released in a given year


```{r}

## Probably that I could do something better today (2018 instead of 2015).

jeux_sortis <- function(year) {
  return(as.data.frame(
    unique(
      data[data$year == year,c(2,6,7)])[order(unique(data[data$year == year,c(2,6,7)])$platform),]
    )	
  )
}

## The example of 2007

head(jeux_sortis(2007))

```


## How many games on each platform in a given year? 

And what percentage of all the games released that year?


```{r}

jeux_proportion <- function(year) { 
  return(rbind(
    sort(table(jeux_sortis(year)$platform)), 
    sort(table(jeux_sortis(year)$platform)) / length(jeux_sortis(year)$platform) * 100)
    )
}

## First line is the number of games
## Second line is the percentage

jeux_proportion(1998)

```


## Table 1: Number of games per platform for all platforms


```{r}

data_t1 <- data
data_t1 <- data_t1[,-c(3,4,5,6,8,10)]
data_t1 <- unique(data_t1)

sort(table(data_t1$platform), decreasing = TRUE)

```


## Figure 1b : number of games per genre and platform

We start by loading informations on platforms: is it a computer, a home console, or a mobile console?

```{r message = FALSE}

## platforms_def.csv contains information on platforms
platf <- read_csv("platforms_def.csv")

head(platf)

```


## Pre-processing of data


```{r message = FALSE}

data_p <- data[data$platform %in% platf$platform,]
data_p <- data_p[data_p$genre != "autre genre",]
data_p <- inner_join(data_p, platf)
data_p <- subset(data_p, subset = data_p$addon != "addon")
data_p <- data_p[!duplicated(data_p[c("ID", "type")]),]
data_p <- data_p[!(data_p$genre == "réflexion"),]
data_p$genre <- factor(data_p$genre)
levels(data_p$genre) <- c("Action", "Action-Adventure", "Adventure", "Role-Playing Game", "Simulation", "Strategy")

```

Now the genres are factors.

## We draw the graph of fig 1b.


```{r}

p2 <- ggplot(data_p, aes(x = genre, fill = type)) +
  scale_fill_grey(labels = c(" PC   ", " Home Console  ", " Mobile Console"),
                  name = element_blank()) +
  geom_bar(position = "dodge") +
  xlab("") +
  ylab("Number of historical videogames\n ") +
  theme(legend.position = "bottom") +
  theme(panel.background = element_rect(fill = "grey92")) +
  theme(axis.ticks = element_blank())

p2

ggsave("fig_1b.png", width = 10)

```

Half of historical videogames in the database are PC strategy games!


## Figure 2a : Historical periods per platforms

The data frame "data.period" will be used to identify the historical period of each game. 

The data frame "data.conflict" will be used to identify the historical conflict for each game. 


```{r message = FALSE}

data <- data[data$platform %in% platf$platform,]
data <- data[data$genre != "autre genre",]
data <- inner_join(data, platf)

data_per <- subset(data, subset = data$addon != "addon")
data_per <- data_per[data_per$period %in% c("préhistoire", "antiquité", "moyen-âge", "epoque moderne", "epoque contemporaine"),]
data_per <- data_per[,c(1, 2, 3, 6, 7, 8, 11)]

data_per2 <- data[!(data$ID %in% data_per$ID),]
data_threshold <- as.data.frame(table(data_per2$period), stringsAsFactors = F)
data_per2 <- data_per2[data_per2$period %in% (data_threshold$Var1[which(data_threshold$Freq >= 3)]),]
colnames(data_per2)[8] <- "event"

per_tab <- read_csv("period.csv", col_names = F)
colnames(per_tab) <- c("event", "period")

data_per2 <- inner_join(data_per2, per_tab)
data_per2 <- data_per2[,c(1, 2, 3, 6, 7, 11, 12)]

hey <- bind_rows(data_per, data_per2)
hey <- unique(hey[,c(1, 2, 3, 4, 6, 7)])

per_plat <- as.data.frame(table(hey$period, hey$type))
colnames(per_plat) <- c("epoque", "platform", "freq")
per_plat$epoque <- ordered(per_plat$epoque, levels = c("préhistoire", "antiquité", "moyen-âge", "epoque moderne", "epoque contemporaine", "toutes époques confondues", "mythes"))
levels(per_plat$epoque) <- c("Prehistory", "Ancient history", "Middle Ages
", "Early modern\nperiod", "Late\nmodern", "All eras", "Myths")

```


## Here we graph figure 2a


```{r}

p3 <- ggplot(per_plat, aes(x = epoque, y = freq, fill = platform)) +
  scale_fill_grey(labels = c(" Computer  ", " Home console  ", " Mobile console"),
                  name = element_blank()) +
  geom_bar(position = "dodge", stat = "identity") +
  xlab("") + ylab("Number of games\n ") +
  theme(legend.position = "bottom") +
  theme(panel.background = element_rect(fill = "grey92")) +
  theme(axis.ticks = element_blank())

p3

ggsave("fig_2a.png", width = 10)

```


## Figure 2b


```{r}

per_type <- as.data.frame(table(hey$genre, hey$period))
per_type <- per_type[!(per_type$Var1 == "réflexion"),]
colnames(per_type) <- c("genre", "epoque", "freq")
per_type$epoque <- ordered(per_type$epoque, levels = c("préhistoire", "antiquité", "moyen-âge", "epoque moderne", "epoque contemporaine", "toutes époques confondues", "mythes"))
levels(per_type$epoque) <- c("Prehistory", "Ancient history", "Middle Ages", "Early modern\nperiod", "Late modern\nperdio", "All eras", "Myths")
per_type$genre <- droplevels(per_type$genre)
levels(per_type$genre) <- c("Action", "Action-Adventure", "Adventure", "Role-Playing Game", "Simulation", "Strategy")

```


```{r}

p4 <- ggplot(per_type, aes(x = epoque, y = freq, fill = genre)) +
  scale_fill_grey(name = element_blank()) +
  geom_bar(position = "dodge", stat = "identity") +
  xlab("") + ylab("Number of games\n ") +
  theme(legend.position = "bottom") +
  theme(panel.background = element_rect(fill = "grey92")) +
  theme(axis.ticks = element_blank())

p4

ggsave("fig_2b.png", width = 10)


```






# Not translated after that. Sorry.

-------------------

```{r}


### Nombre de jeux sur la seconde guerre mondiale
#
# (data$period == "deuxième guerre mondiale") %>% data[.,] %>% as.data.frame %>% .$ID %>% unique %>% length
# (data$period == "front ouest") %>% data[.,] %>% as.data.frame %>% .$ID %>% unique %>% length
# (data$period == "front est") %>% data[.,] %>% as.data.frame %>% .$ID %>% unique %>% length
#
### Nombre de jeux de stratégie sortis sur wii u et ds et 3ds
#
# (data$genre == "stratégie" & data$platform == "nintendo wii u") %>% data[.,] %>% as.data.frame %>% .$ID %>% unique %>% length
# (data$genre == "stratégie" & data$platform == "nintendo ds") %>% data[.,] %>% as.data.frame %>% .$name %>% unique %>% length
# (data$genre == "stratégie" & data$platform == "nintendo ds") %>% data[.,] %>% as.data.frame %>% .$name %>% unique %>% cat(., sep = " ( , ) ")
# (data$genre == "stratégie" & data$platform == "nintendo 3ds") %>% data[.,] %>% as.data.frame %>% .$ID %>% unique %>% length

```



## Les événements historiques selon les années (première sortie d'un jeu)


```{r eval = FALSE}

data.conflict$period <- droplevels(data.conflict$period)

data.add <- data.conflict[data.conflict$ID %in% setdiff(data.conflict$ID, data.period$ID),]
data.add$period <- as.character(data.add$period)
data.add <- data.add[!(data.add$period == "toutes époques confondues"),]

colnames(data.conflict)[grepl("period", colnames(data.conflict))] <- "conflict"

antiquite <- tolower(c("egypte antique", "Guerres médiques", "Conquêtes d'Alexandre le Grand", "Guerres puniques", "Conquêtes romaines", "Trois royaumes", "Invasions barbares"))
moyenage <- tolower(c("invasions Vikings", "Invasions Mongoles", "Croisades", "Guerre d'indépendance écossaise", "Guerre de Cent Ans", "Guerre des Deux Roses"))
moderne <- tolower(c("Conquista", "Sengoku Jidai", "Guerre de Trente Ans", "Guerre du Nord", "Guerre de Succession d'Espagne", "Guerre de Sept Ans", "Guerre indépendance américaine", "Guerres napoléoniennes"))
mythe_ancien <- tolower(c("mythologie amérindienne", "mythologie égyptienne", "mythologie germano-nordique", "mythologie gréco-romaine", "mythologie médiévale", "mythologie orientale", "mythologie russe", "mille et une nuits", "légendes arthuriennes"))
mythe_moderne <- tolower(c("mafia", "western", "piraterie"))

data.add$period[data.add$period %in% antiquite] <- "antiquité"
data.add$period[data.add$period %in% moyenage] <- "moyen-âge"
data.add$period[data.add$period %in% moderne] <- "epoque moderne"
data.add$period[data.add$period %in% mythe_ancien] <- "mythe ancien"
data.add$period[data.add$period %in% mythe_moderne] <- "mythe moderne"

data.add$period[!(data.add$period %in% c("antiquité", "moyen-âge", "epoque moderne", "mythe ancien", "mythe moderne"))] <- "epoque contemporaine"

data.add$period <- factor(data.add$period)

data.period <- rbind(data.period, data.add)

data.period <- data.period[,c(1,2,3,6,8)]

data.conflict <- merge(data.conflict, unique(data.period[,c(1,5)]))

```



