---
title: "mobygames_allgame"
author: "Yannick Rochat"
date: "30/09/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script displays the distribution of videogames released per year on [MobyGames](mobygames.com), an online collaborative database referencing more than a hundred thousand games (in 2018).

We gather the number of games per year by extracting this information from the mobygames summary pages.

```{r}

url <- "http://www.mobygames.com/browse/games/"

```

Choose a time interval.

```{r}

years <- as.character(1971:2017)

```

We extract the numbers and produce a data frame.

```{r}

library(rvest)
library(stringr)

hey <- lapply(years, . %>% 
              str_c(url, ., "/list-games/") 
              %>% read_html() 
              %>% html_nodes(xpath = "//*[@id='main']/div/div[2]/div/text()[3]")
              %>% html_text())

year_games <- unlist(hey)
year_games <- str_extract(year_games, "[0-9]+")
year_games <- as.numeric(year_games)

games <- data.frame(year = as.numeric(years), count = year_games)

```

Let's plot it.

```{r}

library(ggplot2)

g <- ggplot(games, aes(x = year, y = count)) + geom_bar(stat = "identity")
g <- g + ggtitle(str_c("Number of games per year on MobyGames (1971-2017). N = ", sum(games$count), " games"))
g <- g + xlab("Year")
g <- g + ylab("Number of Games")
g

```

And save it.

```{r}

ggsave("mobygames_allgames.png", width = 10)

```
